### 2022-05-19 writing/manuscript.Rmd
revision updates

---
bibliography: themusiclab.bib
csl: apa.csl
header-includes:
- \usepackage[left]{lineno}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{tabu}
- \usepackage{afterpage}
- \usepackage{mdframed}
- \usepackage{xcolor}
- \usepackage{float}
- \definecolor{bleu}{HTML}{2200cc}
notes-after-punctuation: no
output:
  pdf_document:
    fig_caption: yes
    latex_engine: lualatex
urlcolor: bleu
link-citations: yes
linkcolor: bleu
always_allow_html: true
---

# Children infer the behavioral contexts of unfamiliar foreign songs

Courtney B. Hilton$^{1,\wedge,\ast}$, Liam Crowley-de Thierry$^{2,\wedge,\ast}$, Ran Yan$^{1,3}$, Alia Martin$^{2}$, & Samuel A. Mehr$^{1,2,4,\ast}$

@@

$^{1}$Department of Psychology, Harvard University, Cambridge, MA 02138, USA.\
$^{2}$School of Psychology, Victoria University of Wellington, Wellington 6012, New Zealand.\
$^{3}$Department of Psychology, Smith College, Northampton, MA 01063, USA.\
$^{4}$Data Science Initiative, Harvard University, Cambridge, MA 02138, USA.

\*Corresponding author. Emails: [courtneyhilton\@g.harvard.edu](mailto:courtneyhilton@g.harvard.edu){.email}, [crowleliam\@myvuw.ac.nz](mailto:crowleliam@myvuw.ac.nz){.email}, [sam\@wjh.harvard.edu](mailto:sam@wjh.harvard.edu){.email}

**Author Note.** A reproducible version of this manuscript, including all analysis and visualization code and data, is available at <https://github.com/themusiclab/form-function-kids>. Audio excerpts from the *Natural History of Song Discography* are available at <https://osf.io/vcybz> and can be explored interactively at <https://themusiclab.org/nhsplots>; all other study data and materials are available at <https://osf.io/b9tnu>. The study was preregistered at <https://osf.io/56zne>. For assistance, please contact C.B.H., L.C-T., and S.A.M.

@@

```{r libraries, include = FALSE}

library(pacman)
pacman::p_load(lmerTest,
               broom,
               broom.mixed,
               sf,
               colorspace,
               Rmisc,
               ggtext,
               ggpubr,
               kableExtra,
               Hmisc,
               gridExtra,
               scales,
               psycho,
               patchwork,
               png,
               here,
               sjmisc,
               tidyverse,
               scales,
               glue, 
               rsq,
               ggpubr)

@@

# format p-values automatically
format_p <- function(p) {
  if (p < .001) {
    return("< .001")
  } else if (p > .01) {
    return(paste0("= ", round(p, digits = 2)))
  } else {
    return(paste0("= ", round(p, digits = 3)))
  }
}

@@

f <- function(num) {
  round(num,2) %>% format(., nsmall=0, big.mark=",")
}

f2 <- function(num) {
  formatC(num, 2, format="f") %>% format(., nsmall=0, big.mark=",")
}

@@

```{r load_data, include = FALSE}

# checks if preprocessed data exists, if not: runs builder
# (!!you should also re-run script if you have updated the data)
if (!file.exists(here("results", "preprocessed_data.RData"))) { ## SM: for public repo, we'll need to use a de-identified data file as the source (maybe just the post-processing file with any identifiers removed)
  source(here("analysis", "builder.R"))
}

# load all preprocessed data from builder
load(here("results", "preprocessed_data.RData"))

@@

if (!file.exists(here("results", "selected_features.RData"))) { ## SM: for public repo, we'll need to use a de-identified data file as the source (maybe just the post-processing file with any identifiers removed)
  source(here("analysis", "feature_selection.R"))
}

# results from music feature analysis
load(here("results", "selected_features.RData"))

@@

@@

```{r data_lists}

# all data referenced in manuscript will be in one of these lists
kids <- list() # info of kid participants
adults <- list() # info of adult participants
mods <- list() # statistical models
plots <- list() # plots

```{r kid_stats}

# 1 row = 1 kid
kid_participants <- KFC_clean %>%
  group_by(user_id) %>%
  summarise(
    across(c(gender,country,language,age,singOften,playMusicOften), unique),
    gender = if_else(is.na(gender), "not_reported", gender),
    accuracy = mean(correct),
    med_rt = median(rt)
    ) %>% 
  mutate(country = case_when(
    country == "Aotearoa/New Zealand" ~ "New Zealand",
    country == "Kingdom of the Netherlands" ~ "Netherlands",
    country == "Russia" ~ "Russian Federation",
    country == "South Korea" ~ "Republic of Korea",
    TRUE ~ as.character(country)))

# general participant characteristics
kids$n$total <- nrow(kid_participants)
kids$n$countries <- length(unique(kid_participants$country))
kids$n$languages <- length(unique(kid_participants$language))

# compute sample size of adults for Male, Female, Other, and Not reported
for (gend in unique(kid_participants$gender)) {
  kids$n$gender[[{{gend}}]] <- nrow(filter(kid_participants, gender == {{gend}}))
}

# info about which were excluded and why
kids$excl$complete <- exclusions$before_exclusions - exclusions$non_complete
kids$excl$complete_age <- kids$excl$comlete - exclusions$age

# task performance (at participant level)
kids$correct$avg <- mean(kid_participants$accuracy)
kids$correct$sd <- sd(kid_participants$accuracy)

kids$heal$guesses <- KFC_clean %>% 
  filter(type == "HEALING") %>% 
  count(response, sort = TRUE) %>% 
  mutate(pcnt = (n / sum(n)) %>% percent()) %>% 
  select(response, pcnt) %>% 
  pivot_wider(names_from = response, values_from = pcnt)

kids$all$heal_guesses <- KFC_clean %>% 
  filter(response == "HEALING") %>% 
  count(type, sort = TRUE) %>% 
  mutate(pcnt = (n / sum(n)) %>% percent()) %>% 
  select(type, pcnt) %>% 
  pivot_wider(names_from = type, values_from = pcnt)

# proportion of guesses per song type
kids$proportions <- guesses$by_type %>%
  select(-FC) %>%
  group_by(guess) %>%
  summarise(avg = percent(mean(KFC), accuracy = 0.1)) %>%
  split(.$guess)

```{r adult_stats}

# 1 row = 1 adult
adult_participants <- FC_clean %>%
  group_by(user_id) %>%
  summarise(gender = if_else(is.na(unique(gender)), "not_reported", unique(gender)),
            age = unique(age))

# total sample size of adults
adults$n$total <- nrow(adult_participants)

# compute sample size of adults for Male, Female, Other, and Not reported
for (gend in unique(adult_participants$gender)) {
  adults$n$gender[[{{gend}}]] <- nrow(filter(adult_participants, gender == {{gend}}))
}

# mean age
adults$age$avg <- mean(as.numeric(adult_participants$age), na.rm = TRUE)
adults$age$sd <- sd(as.numeric(adult_participants$age), na.rm = TRUE)
adults$age$min <- min(as.numeric(adult_participants$age), na.rm = TRUE)
adults$age$max <- max(as.numeric(adult_participants$age), na.rm = TRUE)

```{r kid_regressions, include = FALSE}

# compute simple linear regression predicting accuracy by "age", "singOften", "playMusicOften"
for (covariate in c("age", "singOften", "playMusicOften")) {
  mods$accuracy[[covariate]] <- lm(accuracy ~ as.numeric(kid_participants[[covariate]]), data = kid_participants) %>% glance()
}

# median RT predicted by age
mods$rt$age <- lm(med_rt ~ age, data = kid_participants) %>% glance()

```{r kids_adults_comparison}

# compute simple linear regression over song types
for (type in guesses$by_type$guess) {
  mods$kfc_fc[[{{type}}]] <- lm(FC ~ KFC, data = filter(guesses$by_type, guess == {{type}})) %>% glance()
}

# Using Kendall's tau
ranked_guesses <- guesses$by_type %>%
  group_by(guess) %>%
  mutate(diff = KFC - FC,
         KFC_rank = as.integer(rank(desc(KFC))),
         FC_rank = as.integer(rank(desc(FC))))

# compute tau correlations over song types
for (type in unique(guesses$by_type$guess)) {
  mods$tau[[{{type}}]] <- cor.test(filter(ranked_guesses, guess == {{type}}) %>% pull(KFC_rank),
                                        filter(ranked_guesses, guess == {{type}}) %>% pull(FC_rank),
                                         method = "kendall") %>% tidy()
}

```{r d-prime_group_level}

for (type_var in c('DANCE', 'LULLABY', 'HEALING')) {
  kids$d$group[[{{str_to_lower(type_var)}}]] <- KFC_clean %>%
    group_by(song, type, response) %>%
    summarise(n = n()) %>%
    mutate(type = if_else(type == type_var, type_var, 'OTHER'),
           response = if_else(response == type_var, type_var, 'OTHER')) %>%
    pivot_wider(names_from = c(type, response), values_from = n, names_sep = "_", values_fn = sum) %>%
    summarise(
      hit = sum(eval(as.name(paste(type_var, type_var, sep = "_")))),
      FA = sum(eval(as.name(paste("OTHER", type_var, sep = "_")))),
      miss = sum(eval(as.name(paste(type_var, "OTHER", sep = "_")))),
      CR = sum(eval(as.name(paste("OTHER", "OTHER", sep = "_"))))) %>%
    replace_na(list(hit = 0, FA = 0, miss = 0, CR = 0)) %>%
    summarise(sensitivity = psycho::dprime(n_hit = sum(hit), n_fa = sum(FA), n_miss = sum(miss), n_cr = sum(CR))) %>%
    summarise(d_prime = sensitivity$dprime, beta = sensitivity$beta, c =  sensitivity$c, type = str_to_lower(type_var))
}

# average d-prime across types
kids$d$group$avg <- bind_rows(kids$d$group) %>% pull(d_prime) %>% mean()

```{r d-prime_participant_level}

# d-primes for each participant for each song type
for (type_var in c('DANCE', 'LULLABY', 'HEALING')) {
  kids$d$ages[[{{str_to_lower(type_var)}}]] <- KFC_clean %>%
    group_by(type, response, age, user_id) %>%
    summarise(n = n()) %>%
    mutate(type = if_else(type == type_var, type_var, 'OTHER'),
           response = if_else(response == type_var, type_var, 'OTHER')) %>%
    pivot_wider(names_from = c(type, response), values_from = n, names_sep = "_", values_fn = sum) %>%
    group_by(age, user_id) %>%
    summarise(
      hit = sum(eval(as.name(paste(type_var, type_var, sep = "_")))),
      FA = sum(eval(as.name(paste("OTHER", type_var, sep = "_")))),
      miss = sum(eval(as.name(paste(type_var, "OTHER", sep = "_")))),
      CR = sum(eval(as.name(paste("OTHER", "OTHER", sep = "_"))))) %>%
    replace_na(list(hit = 0, FA = 0, miss = 0, CR = 0)) %>%
    group_by(age, user_id) %>%
    summarise(sensitivity = psycho::dprime(n_hit = sum(hit), n_fa = sum(FA), n_miss = sum(miss), n_cr = sum(CR))) %>%
    summarise(d_prime = sensitivity$dprime, beta = sensitivity$beta, c =  sensitivity$c, type = str_to_lower(type_var))
}

# d-primes averaged for each age bin
kids$d$avg$ages <- bind_rows(kids$d$ages) %>% #SDT_all_ages
  group_by(type, age) %>%
  summarise(mean_d_prime = mean(d_prime),
            sd_d_prime = sd(d_prime),
            n_kids = n(),
            se_d_prime = sd_d_prime / sqrt(n_kids),
            conf_95 = qt(.975, n_kids - 1) * se_d_prime)

# d-primes averaged for each song type
kids$d$avg$types <- bind_rows(kids$d$ages) %>%
  group_by(type) %>%
  summarise(mean_d_prime = mean(d_prime),
            sd_d_prime = sd(d_prime),
            n_kids = n(),
            se_d_prime = sd_d_prime / sqrt(n_kids),
            conf_95 = qt(.975, n_kids - 1) * se_d_prime)

```{r d-prime_regressions}

# t.tests + simple linear models
for (type_x in c("dance", "healing", "lullaby")) {
  mods$d$t.test[[{{type_x}}]] <- t.test(kids$d$ages[[{{type_x}}]]$d_prime, mu = 0, alternative = "two.sided")
  mods$d$lm[[{{type_x}}]] <- lm(d_prime ~ age, data = kids$d$ages[[{{type_x}}]]) %>% tidy()
}

# linear model across all ages for all song types
mods$d$lm$all <- bind_rows(kids$d$ages) %>% 
  ungroup() %>% 
  mutate(age = center(age)) %>% 
  lmer(d_prime ~ age + (1|user_id) + (type|age), data = .) %>% anova() %>% tidy()

# add d-primes to participant list
kids$d$all <- kid_participants %>%
  inner_join(.,
             select(bind_rows(kids$d$ages), user_id, d_prime, type),
             by = "user_id")

# linear regression on accuracy by parent singing to child + playinng recorded music
mods$d$parentSing <- lm(d_prime ~ as.numeric(singOften), data = kids$d$all) %>% glance()
mods$d$playMusic <- lm(d_prime ~ as.numeric(playMusicOften), data = kids$d$all) %>% glance()

```{r correlations_between_KFC_FC_guesses}

# compute correlations between KFC and FC over ages
kids$corr$wFC <- guesses$by_type_age %>%
  group_by(age, response) %>%
  summarise(KFC_FC_correlation = cor(FC, KFC))

# linear regression on R^2 of child-adult guessing correlations by child age
mods$kfc_fc$corr$age <- lm(KFC_FC_correlation ~ age, data = kids$corr$wFC ) %>% glance()

# testing whether multiple regression models for kids and adults are statistically equivalent
for (type_x in c("dance", "lullaby", "healing")) {
  # pool + wrangle data
  tmp <- songs %>% 
    pivot_longer(cols = c(glue("KFC_{type_x}"), glue("FC_{type_x}")), names_to = type_x, values_to = "probs")

  # define model formulae
  x1 <- reformulate(selected_features[[type_x]], response = "probs")
  x2 <- reformulate(c(selected_features[[type_x]], type_x), response = "probs")

  # run models
  m1 <- lm(x1, data = tmp)
  m2 <- lm(x2, data = tmp)

  # test using Chi-square test
  mods$equiv[[type_x]] <- anova(m1,m2, test = "Chisq") %>% tidy() %>% slice(., 2)
}

```{r multi-regression_models}

# Compute multiple regression models on LASSO selected features
for (data_x in c("KFC", "FC")) {
  for (type_x in c("dance", "lullaby", "healing")) {
    mods$feature[[data_x]][[type_x]] <- lm(
      reformulate(selected_features[[type_x]], response = paste(data_x, type_x, sep = "_")),
      data = songs)

    mods$feature$s[[data_x]][[type_x]] <- mods$feature[[data_x]][[type_x]] %>% glance()
  }
}

```{=tex}
\normalsize
\begin{mdframed}[backgroundcolor=gray!20]
Music commonly appears in a variety of behavioral contexts in which it can be seen as playing a functional role, such as when a parent sings a song with the goal of soothing their child. Humans readily make inferences about such behavioral contexts based on sound alone. These inferences tend to be accurate, even if the songs are in foreign languages or unfamiliar musical idioms: upon hearing a Blackfoot lullaby, a Korean listener with no experience of Blackfoot music, language, or broader culture is far more likely to judge the music’s function as “to soothe a baby” than as “for dancing”. Are such inferences shaped by musical exposure or does the human mind naturally detect links between musical form and function? Children's developing experiences with music provide a clear test of this question. We studied musical inferences in a large sample of children ($N$  = `r kids$n$total %>% f`), who heard dance, lullaby, and healing songs from `r length(unique(songs$culture))` world cultures and who were tasked with guessing the original behavioral context in which each was performed. We found little evidence for an effect of experience: children reliably inferred the original behavioral contexts with only minimal improvement in performance from the youngest (age 3 or younger) to the oldest (age 12). Children's inferences tightly correlated with those of adults for the same songs, as collected from a similar massive online experiment ($N$ = `r adults$n$total %>% f`). Moreover, the same musical/acoustical features were predictive of the inferences of both samples. These findings suggest that accurate inferences about the behavioral contexts of music, driven by universal links between form and function in music across cultures, do not require extensive musical experience.

@@

\linenumbers

Music is a ubiquitous part of human culture that plays a variety of functional roles in our day-to-day lives [@Nettl2005; @Mehr2019; @Trehub2015a; @Lomax1968]: We sing songs to entertain each other, to find spiritual connection, to tell stories, to regulate our emotions, or just to pass the time.

Whereas some musical contexts are idiosyncratic, like how complex geographical knowledge is culturally encoded in song by some groups of Indigenous Australians [@Norris2014], others are recurrent across societies. These latter contexts tend to also be accompanied by recurrent sets of features. For example, in Western societies, lullabies tend to have slow tempos; smooth, minimally accented contours; and a narrow dynamic and pitch range [@Trehub1993; @Trehub1998; @Bergeson1999], a pattern replicated in a globally representative sample of societies [@Mehr2019]. That music appears universally in conjunction with a given behavioral context (e.g., infant care), and that examples of music in that context share particular musical features (e.g., minimal accents), implies a link between form and function in music production.

Adults are sensitive to this link, in that they reliably distinguish song functions even when examples are unfamiliar and foreign. This has been demonstrated in experiments where naïve listeners are asked to identify the context in which a song was originally used, purely on the basis of the sounds it contains. Indeed, they successfully identify lullabies when paired with love songs as foils [@Trehub1993a] and distinguish lullabies, dance songs, healing songs, and love songs from one another [@Mehr2018a; @Mehr2019].

How do adults come to be sensitive to these musical form-function links? Across the lifespan, musical experience is ubiquitous and rich, including through infancy [@Mendoza2021] and childhood [@Bonneville-Roussy2013], so it's plausible that a sensitivity to form and function in music is shaped by direct musical experience. This is especially so given people's detailed long-term memories for music [@Levitin1996; @Krumhansl2010a] and the importance of implicit associative learning in musical knowledge [@Rohrmeier2012].

Two forms of evidence point to an alternative explanation, however. First, systematic form-function pairings are found in vocalizations throughout the animal kingdom [@Morton1977; @Fitch2002], and likely reflect innate aspects of vocal signaling [@Darwin1871; @Mehr2020]. Indeed, emotional vocalizations are not only cross-culturally intelligible within humans [@Scherer2001; @Chronaki2018; including in music: @Balkwill1999], but also between species [@Filippi2017]. Humans extend these emotional inferences to environmental sounds [@Ma2015] and even to abstract shapes [@Sievers2019]. Form-function inferences in music might borrow from and extend this repertoire of instinctive signaling mechanisms.

Second, despite their lesser degree of exposure to music, relative to adults, infants are affected by music in ways that are predictable from the behavioral functions of music. For example, infants relax more in response to unfamiliar foreign lullabies than non-lullabies [@Bainbridge2021], and move rhythmically more in response to metrical music than speech [@Zentner2010]. While neither implies that infants are *aware* of the function of a song, early-developing responses to music may eventually support form-function inferences when combined with precocial social inferences [@Powell2013; @Bohn2019; @Mehr2016]; biases for socially meaningful stimuli [@Frank2014]; and functional intuitions about other socially embedded vocalizations, such as speech [@Martin2012; @Vouloumanos2014].

To test between these perspectives, we studied musical inferences across early and middle childhood, focusing on whether and how these inferences change as children's musical experience grows. Children participated in a citizen-science experiment (readers can try it with their children at <https://themusiclab.org>) where they listened to examples of lullabies, dance songs, and healing songs from `r length(unique(songs$culture))` mostly small-scale societies, drawn from the *Natural History of Song Discography* [@Mehr2019], and were tasked with guessing the original behavioral context.

@@

Children visited the citizen science website <https://themusiclab.org> to participate. We posted links to the experiment on social media, which were spread via organic sharing, and we also advertised the experiment on <https://childrenhelpingscience.com>, a website disseminating information on web-based research for children [@Sheskin2020]. Recruitment was/is open-ended; at the time of analysis, we had complete data from `r kids$excl$complete %>% f` children with reported ages from "3 or under" to "17 or older". Given our interest in early development, we chose to analyze only data from the youngest 10 age bands (ages "3 or under" to "12"). We excluded older children ($n$ = `r exclusions$age %>% f`); children whose parents indicated that they had assisted their child during the experiment ($n$ = `r exclusions$parentHelp %>% f`); children who indicated they had played the game previously ($n$ = `r exclusions$playedBefore %>% f`); children with known hearing impairments ($n$ = `r exclusions$hearing %>% f`); and children who were missing data from the test phase ($n$ = `r exclusions$missing %>% f`).

This left `r kids$n$total %>% f` participants for analysis (`r kids$n$gender$Male %>% f` male, `r kids$n$gender$Female %>% f` female, `r kids$n$gender$Other %>% f` other; the cohort's breakdown of age and gender, after exclusions, is in SI Figure 1) representing `r kids$n$languages %>% f` native languages and `r kids$n$countries %>% f` countries (SI Tables 1 and 2).

No compensation was given, reducing the likelihood that adults were posing as children, but as an additional check, we examined participants' response times during the study. Consistent with prior work showing that response time in perceptual and cognitive tasks decreases over the course of childhood [@Hale1990; @Kiselev2009], response times during the task decreased reliably as a function of age (*F*(1, `r mods$rt$age$df.residual %>% f`) = `r mods$rt$age$statistic %>% f`, *p* `r mods$rt$age$p.value %>% format_p()`, *R*^2^ `r mods$rt$age$r.squared %>% format_p()`; SI Figure 2). More generally, there is growing evidence that online and lab-based data collection produce similar results, with limited evidence for reduced reliability in online cohorts [@Germine2012; @Hartshorne2019].

@@

Song excerpts were drawn from the *Natural History of Song Discography* [@Mehr2019], a corpus of vocal music collected from 86 mostly small-scale societies, including hunter-gatherers, pastoralists, and subsistence farmers. Each song in the corpus was originally performed in one of four behavioral contexts: dance, healing, lullaby, and love. All recordings were selected on the basis of supporting ethnographic material, as opposed to the acoustic features of the songs [see: @Mehr2019; @Mehr2018a].

@@

```{r fig1, fig.align = "center", fig.height = 7.5, fig.width = 12, out.extra = 'trim={0 2.5cm 0 2.5cm}, clip', fig.cap = "\\textbf{Figure 1 | Schematic of the test phase of the experiment.} Children participated in an online game, narrated by \\textit{Susie the Star}, whose instructions were spoken aloud and printed in speech bubbles on the screen. Parents helped their children get accustomed to the procedure, but were asked not to help during the test phase of the experiment. During testing (panels \\textbf{a-c}) children listened to a series of songs; after each one, \\textit{Susie} asked them to guess its behavioral context."}
experiment_schematic <- readPNG(here("viz", "kfc_procedureSchematic.png"))
plot.new()
rasterImage(experiment_schematic,0,0,1,1)
```

@@

# Context

>>>>

---
bibliography: themusiclab.bib
csl: apa.csl
header-includes:
- \usepackage{ragged2e}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{tabu}
- \usepackage{afterpage}
- \usepackage{mdframed}
- \usepackage{xcolor}
- \usepackage{soul}
- \usepackage{float}
- \definecolor{bleu}{HTML}{2200cc}
notes-after-punctuation: no
output:
  pdf_document: 
    fig_caption: yes
    latex_engine: xelatex
  word_document: default
urlcolor: bleu
link-citations: yes
linkcolor: bleu
always_allow_html: yes
---

\raggedright
\LARGE
\textbf{Children infer the behavioral contexts of unfamiliar foreign songs}

\vspace{0.1in}

\justifying
\normalsize
Courtney B. Hilton$^{1,2,\wedge,\ast}$, Liam Crowley-de Thierry$^{3,\wedge,\ast}$, Ran Yan$^{2,4}$, Alia Martin$^{3}$, & Samuel A. Mehr$^{1,2,5,\ast}$

\footnotesize
$^{1}$Haskins Laboratories, New Haven, CT 06511, USA.\
$^{2}$Department of Psychology, Harvard University, Cambridge, MA 02138, USA.\
$^{3}$School of Psychology, Victoria University of Wellington, Wellington 6012, New Zealand.\
$^{4}$Department of Psychology, University of Michigan, Ann Arbor, MI 48109, USA.\
$^{5}$Data Science Initiative, Harvard University, Cambridge, MA 02138, USA.

\*Corresponding author. Emails: [courtneyhilton\@g.harvard.edu](mailto:courtneyhilton@g.harvard.edu){.email}, [liam.crowleydethierry\@vuw.ac.nz](mailto:liam.crowleydethierry@vuw.ac.nz){.email}, [sam\@wjh.harvard.edu](mailto:sam@wjh.harvard.edu){.email}

@@

**Author Note.** All data and materials, including a reproducible version of this manuscript with all analysis and visualization code, are available at <https://github.com/themusiclab/form-function-kids>. Audio excerpts from the *Natural History of Song Discography* are available at <https://osf.io/vcybz> and can be explored interactively at <https://themusiclab.org/nhsplots>. The study was preregistered at <https://osf.io/56zne>. For assistance, please contact C.B.H., L.C-T., and S.A.M.

@@

```{r libraries}

library(pacman)
p_load(
  lmerTest,
  broom,
  broom.mixed,
  sf,
  colorspace,
  Rmisc,
  ggtext,
  ggpubr,
  kableExtra,
  Hmisc,
  gridExtra,
  scales,
  psycho,
  patchwork,
  png,
  here,
  sjmisc,
  tidyverse,
  scales,
  glue, 
  rsq,
  ggpubr
)

@@

# format p-values automatically
fp <- function(p) {
  if (p < .001) {
    return("< .001")
  } else if (p > .01) {
    return(paste0("= ", round(p, digits = 2)))
  } else {
    return(paste0("= ", round(p, digits = 3)))
  }
}

@@

f <- function(num) {format(num, big.mark=",")}
r1 <- function(num) {round(num,1) %>% format(., nsmall=1, big.mark=",")}
r2 <- function(num) {round(num,2) %>% format(., nsmall=2, big.mark=",")}
r3 <- function(num) {round(num,3) %>% format(., nsmall=3, big.mark=",")}
p1 <- function(num) {percent(num, accuracy = 0.1)}
p0 <- function(num) {percent(num, accuracy = 1)}

@@

```{r load_data}

####################################################################
# This manuscript contains computations that take a long time to run
# For convenience, these computations are parameterized such that
# they are not run each time the manuscript is knitted by default
# and instead just load the last saved run.
# The same is true for figures, which do not need to be regenerated
# each time the manuscript is knitted.
# If you want to regenerate anything in your knit, set full_run
# below to TRUE. Otherwise, do nothing.
####################################################################

full_run <- FALSE

if (full_run == TRUE) {
  source(here("analysis", "builder.R"))
  source(here("analysis", "feature_selection.R"))
  source(here("analysis", "analyses.R"))
  knitr::knit(here("viz", "figures.Rmd"))
} else {
  load(here("results", "preprocessed_data.RData"))
  load(here("results", "lasso_analyses.RData"))
  load(here("results", "analyses.RData"))
}

@@

@@

adult_covariates <- read_csv(here('data', "covariatesFC.csv")) %>%
  group_by(user_id) %>%
  summarise(gender = if_else(is.na(unique(gender)), "not_reported", unique(gender)),
            age = unique(age),
            country = unique(country),
            language = unique(language))

@@

```{=tex}
\normalsize
\begin{mdframed}[backgroundcolor=gray!20]
Music commonly appears in behavioral contexts in which it can be seen as playing a functional role, as when a parent sings a lullaby with the goal of soothing a baby. Humans readily make inferences, on the basis of the sounds they hear, regarding the behavioral contexts associated with music. These inferences tend to be accurate, even if the songs are in foreign languages or unfamiliar musical idioms; upon hearing a Blackfoot lullaby, a Korean listener with no experience of Blackfoot music, language, or broader culture is far more likely to judge the music’s function as ``used to soothe a baby" than ``used for dancing". Are such inferences shaped by musical exposure or does the human mind naturally detect links between these kinds of musical form and function? Children's developing experience of music provides a clear test of this question. We studied musical inferences in a large sample of children recruited online ($N$  = `r kids$n$total %>% f`), who heard dance, lullaby, and healing songs from `r length(unique(songs$culture))` world cultures and who were tasked with guessing the original behavioral context in which each was performed. Children reliably inferred the original behavioral contexts with only minimal improvement in performance from the youngest (age 4) to the oldest (age 16), providing little evidence for an effect of experience. Children's inferences tightly correlated with those of adults for the same songs, as collected from a similar online experiment ($N$ = `r adults$n$total %>% f`). Moreover, similar acoustical features were predictive of the inferences of both samples. These findings suggest that accurate inferences about the behavioral contexts of music, driven by universal links between form and function in music across cultures, do not always require extensive musical experience.

@@

Music is a ubiquitous part of human culture that plays a variety of functional roles in our day-to-day lives [@Mehr2019; @Trehub2015a; @Nettl2005; @Lomax1968]: We sing songs to entertain each other, to find spiritual connection, to tell stories, to regulate our emotions, or just to pass the time.

Whereas some musical contexts are idiosyncratic, as in the encoding of complex geographical knowledge in song by some groups of Indigenous Australians [@Norris2014], other contexts are recurrent across societies and tend to co-occur with the same acoustic features. For example, lullabies tend to have slow tempos; smooth, minimally accented contours; and a narrow dynamic and pitch range, as has been long-established in western societies [@Bergeson1999; @Trehub1998a; @Trehub1993], and recently in more globally representative samples of societies [@Hilton2021a; @Mehr2019]. The findings that music appears universally in conjunction with particular behavioral contexts (e.g., infant care), and that examples of music in a given context share particular acoustic features (e.g., minimal accents), imply a link between form and function in music production.

Adults are sensitive to this link, in that they reliably distinguish song functions even when examples are unfamiliar and foreign. This has been demonstrated in experiments where naïve listeners are asked to identify the context in which a song was originally used, purely on the basis of the sounds it contains. Adult listeners indeed identify lullabies when paired with love songs as foils [@Trehub1993a] and distinguish lullabies, dance songs, healing songs, and love songs from one another [@Mehr2019; @Mehr2018a].

How do adults come to be sensitive to these musical form-function links? Across the lifespan, musical experience is ubiquitous and rich, including through infancy [@Mendoza2021; @Yan2021] and childhood [@Bonneville-Roussy2013; @Mehr2014]. It is therefore plausible that a sensitivity to form and function in music is the result of acquired associations through direct musical experience; especially so given humans' detailed long-term memories for music [@Krumhansl2010a; @Levitin1996], including in infancy [@Mehr2016; @Saffran2000], and the importance of implicit associative learning in musical knowledge [@Rohrmeier2012].

Two forms of evidence point to an alternative explanation, however. First, systematic form-function pairings are found in vocalizations throughout the animal kingdom [@Fitch2002; @Morton1977], and are thought to reflect innate aspects of vocal signaling [@Mehr2021; @Darwin1871]. Indeed, emotional vocalizations are not only cross-culturally intelligible within humans [in speech: @Cowen2019; @Chronaki2018; @Scherer2001; in music: @Balkwill1999], but also between species [@Kamiloglu2020; @Filippi2017]. Humans even extend these emotional inferences to non-animate environmental sounds [@Ma2015] and abstract shapes [@Sievers2019]. And a broader class of innate psychoacoustical relationships are thought to underlie the surprisingly robust cross-linguistic associations between certain speech sounds and meaning in the world's languages [@Yu2021; @Blasi2016; @Imai2008; @Nuckolls1999]. Form-function inferences in music may rely upon similar, largely-innate mechanisms.

Second, despite their lesser degree of exposure to music, relative to adults, infants are affected by music in ways that are predictable from behavioral function. For example, infants relax more in response to lullabies than non-lullabies, even when the songs are in unfamiliar languages and from unfamiliar cultures [@Bainbridge2021]; and infants move rhythmically more in response to metrical music than speech [@Zentner2010]. While neither implies that infants are *aware* of the functions of the songs they hear, these early-developing responses to music may provide a grounding for later higher-level social inferences about music, which may develop through interaction with precocial social intelligence [@Bohn2019; @Mehr2016; @Mehr2017; @Powell2013; @Herrmann2007; @Tomasello2005], along with functional intuitions about other social signals, such as speech [@Vouloumanos2014; @Martin2012].

To test between these competing explanations — invoking learned vs. innate mechanisms — we studied musical inferences across early and middle childhood, focusing on whether and how these inferences change as children's musical experience grows. Children participated in a citizen-science experiment (readers can try it with their children at <https://themusiclab.org/quizzes/kfc>) where they listened to examples of lullabies, dance songs, and healing songs from `r length(unique(songs$culture))` mostly small-scale societies, drawn from the *Natural History of Song Discography* [@Mehr2019], and were tasked with guessing the original behavioral context.

@@

Children visited the citizen science website <https://themusiclab.org> to participate. We posted links to the experiment on social media, which were spread via organic sharing, and we also advertised the experiment on <https://childrenhelpingscience.com>, a website disseminating information on web-based research for children [@Sheskin2020]. Recruitment was/is open-ended; at the time of analysis, we had complete data from `r kids$excl$complete %>% f` participants with reported ages from "3 or under" to "17 or older" in one-year increments.

We opted to only analyze the precisely year-long age bands, excluding the participants who responded with "3 or under" and "17 or older" ($n$ = `r exclusions$age %>% f`) since these are harder to interpret (but see SI Text 1 for a supplementary analysis of these participants). We also excluded children whose parents indicated that they had assisted their child during the experiment ($n$ = `r exclusions$parentHelp %>% f`); children who indicated they had played the game previously ($n$ = `r exclusions$playedBefore %>% f`); children with known hearing impairments ($n$ = `r exclusions$hearing %>% f`); children who were missing data from the test phase ($n$ = `r exclusions$missing %>% f`); and children who had trials with responses faster than 100ms or slower than 10 seconds ($n$ = `r exclusions$rt %>% f`). While the exclusion criteria for response times were not preregistered, we note that the main findings replicate under a variety of alternative exclusion decisions, or when not excluding any trials at all.

This left `r kids$n$total %>% f` participants for analysis (`r kids$n$gender$Male %>% f` male, `r kids$n$gender$Female %>% f` female, `r kids$n$gender$Other %>% f` other; the cohort's breakdown of age and gender, after exclusions, is in SI Figure 1). Participants reported speaking `r kids$n$languages %>% f` native languages and being located in `r kids$n$countries %>% f` countries (SI Tables 1 and 2).

Readers should note that the experiment itself was conducted in English (including both written and spoken/audible instructions), however. As such, despite the apparent diversity of countries of origin and native languages, these data should not be considered a representative sample of cultural or linguistic experiences, and is likely to be biased in certain ways (e.g., those with internet access who can speak English; see Discussion).

@@

Song excerpts were drawn from the *Natural History of Song Discography* [@Mehr2019], a corpus of vocal music collected from 86 mostly small-scale societies, including hunter-gatherers, pastoralists, and subsistence farmers. Each song in this corpus was originally performed in one of four behavioral contexts: dance, healing, lullaby, and love. All recordings were selected on the basis of supporting ethnographic material, as opposed to the acoustic features of the songs [see: @Mehr2018a; @Mehr2019].

@@

```{r fig1}
#| fig.cap = "\\textbf{Figure 1.} Schematic of the test phase of the experiment. Children listened to the songs and indicated their responses in the context of an online game. \\textbf{a}, Parents guided their children's participation with the assistance of a progress bar, which indicated which portions of the experiment were intended to be for parents alone, parents and children together, or children alone. The progress bar was visible throughout the experiment. \\textbf{b}, The game was narrated by \\textit{Susie the Star}, an animated character whose instructions were spoken aloud and printed in speech bubbles on the screen. \\textbf{c}, During testing children listened to a series of songs. \\textbf{d}, After each song, \\textit{Susie} asked them to guess its behavioral context and indicate their response by clicking or tapping an image (with accompanying text)."
knitr::include_graphics(here("viz", "figures", "fig1-1-mod.pdf"))
```

@@

# Context of the research
